### 2021-09-28 main.tex
updates

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{soul}
\usepackage{algorithm2e}
\usepackage{bm}
% \usepackage{setspace}
% \linespread{1.8}

\newtheorem{prop}{Proposition}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\bruno}[1]{\textcolor{blue}{#1}} % Bruno Sanso edits - blue
\newcommand{\makenote}[1]{\textcolor{red}{#1}} % Peter edits - red
\newcommand{\findcite}{\textcolor{red}{[Find Citation]}}
\newcommand{\needcite}{\findcite}
\newcommand{\Chi}{\mbox{\Large$\chi$}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\inorm}[1]{\norm{#1}_{\infty}}
\newcommand{\pnorm}[2]{\norm{#1}_{#2}}

\title{Inference for multivariate peaks over threshold models}
\author{Peter Trubey and Bruno Sans\'o\\
Department of Statistics, University of California Santa Cruz\\
{\tt ptrubey@ucsc.edu, bruno@soe.ucsc.edu}}

\begin{document}

%--------title page
\maketitle
\thispagestyle{empty}
\begin{abstract}
  \input{abstract}
\end{abstract}

%\newpage
%--------Table of contents
%\tableofcontents
%\thispagestyle{empty}

%\newpage

%--------content
\input{introduction}

\input{methodology}

\input{evaluation}

\input{results}

\input{conclusion}

\bibliographystyle{chicago}
\bibliography{./refs}

\end{document}

% EOF