### 2021-11-11 evaluation.tex
updates

\section{Scoring criteria for distributions in ${\mathbb S}_\infty^{d-1}$\label{sec:evaluation}}

In order to assess and compare the estimation of a distribution on ${\mathbb S}_\infty^{d-1}$ we consider the theory of proper scoring rules developed in \cite{gneiting2007}.
As mentioned in Section~\ref{subsec:projgamma}, our approach does not provide a density on ${\mathbb S}_{\infty}^{d-1}$, restricting our ability to construct model selection criteria to sample-based approaches.  We consider two scoring methods:  \emph{posterior predictive loss} (PPL) introduced in \cite{gelfand1998}, and \emph{energy score} (ES) introduced in \cite{gneiting2007}.

\subsection{Posterior predictive loss}
The posterior predictive loss criterion is introduced in \cite{gelfand1998}.  When we
  assume a squared error loss function, then for the $i$th observation, the posterior predictive
  loss criterion is computed as
  \begin{equation}
    \label{eq:ppl}
      S_{\text{PPL}}^k\left(P,x_i\right) = \text{Var}_P\left(X_i\right) + \frac{k}{k+1}\left(\text{E}_P\left[X_{i}\right] - x_i\right)^2
  \end{equation}
  where $X_i$ is a random variable from the posterior predictive distribution for $x_i$.  The
  scalar $k$ is a weighting factor that scales the importance of goodness of fit
  relative to precision.  In fact a small $\text{Var}(X_{i})$ indicates high
  precision, and a small $(\text{E}[X_{i}] - x_{i})^2$  indicates a good fit.  In our analysis,
  we take the limit as $k\to\infty$, weighting both
  equally. Overall, the smaller the score, the better.  Note that the PPLC score is defined 
  for a univariate distribution.  As we are dealing
  with a multivariate distribution, we will be taking the posterior predictive loss
  summed over all dimensions.  That is,
  \begin{equation}
    \label{eq:ppl2}
      S_{\text{PPL}}\left(P,\bm{x}_i\right) = \frac{1}{d}\sum_{\ell = 1}^d\left[\text{Var}_P\left(X_{i\ell}\right) + \left(\text{E}_P\left[X_{i\ell}\right] - x_{i\ell}\right)^2\right] .
  \end{equation}
  We report the average $S_{\text{PPL}}$ taken from all observed data.

\subsection{Energy score}
We focus on \emph{energy scores} defined for a general probability distribution $P$, with finite expectation, as 
    \begin{equation}
    \label{eq:es}
    \text{S}_{\text{ES}}\left(P, \bm{x}_i\right) =  \text{E}_p g\left(\bm{X}_i, \bm{x}_i\right) -
                                    \frac{1}{2}\text{E}_p g\left(\bm{X}_i,\bm{X}_i^{\prime}\right),
    \end{equation}
where $g$ is a kernel function. The score defined in Equation \eqref{eq:es} can be evaluated
  using samples from $P$, with the help of the law of large numbers.
  Moreover, Theorem~4 in \cite{gneiting2007}, states that if $g(\cdot,\cdot)$ is a negative 
  definite kernel, then $\text{S}(P,\bm{x})$ is a \emph{proper} scoring rule.  Recall that a 
  real valued function $g$ is a negative definite kernel if it is symmetric in its arguments, 
  and $\sum_{i=1}^n\sum_{j=1}^na_ia_jg(x_i,x_j)\leq 0$ for all positive integers $n$, and any 
  collection $a_1,\ldots,a_n\in{\mathbb R}$ such that  $\sum_{i = 1}^na_i = 0$.

In a Euclidean space, these conditions are satisfied by the Euclidean distance. However, in
  ${\mathbb S}_{\infty}^{d-1}$, Euclidean distance will tend to under-estimate the actual 
  distance required to travel between two points.  Let
\begin{equation*}
    {\mathbb C}_{\ell}^{d-1} = \lbrace \bm{x} : \bm{x} \in {\mathbb S}_{\infty}^{d-1}, x_{\ell} = 1\rbrace
\end{equation*}
comprise the $\ell$th \emph{face}.  For points on the same face, Euclidean distance 
  corresponds to the length of the shortest possible path in ${\mathbb S}_{\infty}^{d-1}$.  
  For points on different faces, the Euclidean distance is a lower bound for that length.

For a finite $p$, the shortest connecting path between two points in ${\mathbb S}_p^{d-1}$
  is the minimum geodesic; its length satisfying the definition of a distance.  Thus its
  length can be used as a negative definite kernel for the purpose of defining an energy
  score. Unfortunately as $p\to\infty$ the resulting surface ${\mathbb S}_{\infty}^{d-1}$
  is not differentiable, implying that routines to calculate geodesics are not readily
  available.  However, as ${\mathbb S}_{\infty}^{d-1}$ is a portion of of a $d$-cube, we
  can borrow a result from geometry \citep{pappas1989} stating that the length of the
  shortest path between two points on a geometric figure corresponds to the length of a
  straight line drawn between the points on an appropriate unfolding, rotation, or \emph{net} of the figure from a $d$-dimensional to a $d-1$-dimensional space.  The optimal net will have the shortest straight line between the points, as long as that line is fully contained within such net. As ${\mathbb S}_{\infty}^{d-1}$ has $d$ faces---each face pairwise adjacent, there are $d!$ possible nets.  However, we are only interested in nets that begin and end on the source and destination faces respectively.  That reduces the number of nets under consideration to $\sum_{k = 0}^{d-2}\binom{d-2}{k}$.  This is still computationally burdensome for a large number of dimensions.

To reduce the computations needed to calculate the energy score we define a kernel as {\bf Please write a clearer description here.} our kernel an upper bound on distance between points on separate faces as the length of the path from the source point, to an optimal point on the intersection between faces, to the destination point.
\begin{prop}\label{prop:g}
Let $\bm{a} \in {\mathbb C}_{\ell}^{d-1}$, and $\bm{b} \in {\mathbb C}_{\jmath}^{d-1}$, for $\ell, \jmath \in \{1, \ldots , d\}$. Define $g$ as
\[  
    g(\bm{a},\bm{b}) = \min_{\bm{c} \in{\mathbb C}_{\jmath}^{d-1}\cap{\mathbb C}_{\ell}^{d-1}}\left\{ 
        \pnorm{\bm{c}-\bm{a}}{2} + \pnorm{\bm{b}-\bm{c}}{2} \right\}.
\]
Then $g$  is negative definite kernel.
\end{prop}

{\em Proof:}
When $\bm{a}$ and $\bm{b}$ are on the same face, then $\ell = \jmath$, and $g$ is the Euclidean distance between $\bm{a}$ and $\bm{b}$, which is a  negative definite kernel.  When $\bm{a}$ and $\bm{b}$ are on separate
  faces, let 
  \[\bm{d} = 
    \argmin_{\bm{c} \in{\mathbb C}_{\jmath}^{d-1}\cap{\mathbb C}_{\ell}^{d-1}}\left\{ 
        \pnorm{\bm{c}-\bm{a}}{2} + \pnorm{\bm{b}-\bm{c}}{2} \right\}.
  \]
  We have that $g(\bm{a}, \bm{b}) = g(\bm{b}, \bm{a})$ due to the uniqueness of $\bm{d}$ and the symmetry of the Euclidean distance. Additionally, given $n$ and  a collection $\alpha_1,\ldots,\alpha_n\in{\mathbb R}$ such that  $\sum_{i = 1}^n\alpha_i = 0$, then, 
  \begin{equation*}
    \begin{aligned}
      \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j g(\bm{x}_1,\bm{x}_2) &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j \bigg[\pnorm{\bm{d} - \bm{x}_1}{2} + \pnorm{\bm{x}_2 - \bm{d}}{2}\bigg]\\
      &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{d} - \bm{x}_1}{2} + \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{x}_2 - \bm{d}}{2} \leq0,
    \end{aligned}
  \end{equation*}
  as the Euclidean norm is a negative definite kernel.$\hfill\Box$

Evaluating $g$ as defined in Proposition \ref{prop:g} requires the solution of a $(d-2)$-dimensional 
  optimization problem.  The following proposition provides a straightforward solution.
\begin{prop}
    Let $\bm{a} \in {\mathbb C}_{\ell}^{d-1}$, and $\bm{b} \in {\mathbb C}_{\jmath}^{d-1}$, for $\ell, \jmath \in \{1, \ldots , d\}$.  For $\ell\neq \jmath$, the 
  transformation $P_{\jmath\ell}(\cdot)$ required to rotate the $\jmath$th face along the $\ell$th axis 
  produces the vector $\bm{b}^\prime$, where
  \begin{equation}
    \label{eqn:rotation}
     \bm{b}^{\prime}_i = P_{\jmath\ell}(\bm{b}) = 
    \begin{cases}
        b_{i} &\text{for }i\neq \jmath,\ell\\
        1 &\text{for }i = \ell\\
        2 - b_{\ell} &\text{for }i = \jmath
    \end{cases}.
  \end{equation}
  Then $g(\bm{a},\bm{b}) = \pnorm{\bm{a} - \bm{b}^{\prime}}{2}$.
\end{prop}
{\em Proof:}
Notice that for 
  $\bm{c} \in {\mathbb C}_{\jmath}^{d-1}\cap{\mathbb C}_{\ell}^{d-1}$,
     $\pnorm{\bm{b} - \bm{c}}{2} =  \pnorm{\bm{b}^{\prime} - \bm{c}}{2}$. {\bf This is true just because you do the math, the invariance of Euclidean norms is not needed.} Then, recalling the definition of $\bm{d}$ in the proof of Proposition \ref{prop:g}, we have that
  \begin{equation*}
    \begin{aligned}
    g(\bm{a}, \bm{b}) &= \pnorm{\bm{a} - \bm{d}}{2} + \pnorm{\bm{b} - \bm{d}}{2}\\
    &= \pnorm{\bm{a} - \bm{d}}{2} + \pnorm{\bm{b}^{\prime} - \bm{d}}{2} = \pnorm{\bm{a} - \bm{b}^{\prime}}{2}.
    \end{aligned}
  \end{equation*}
  The last equality is due to the fact that $\bm{a}$ and $\bm{b}'$ belong to the same hyperplane.
  $\hfill\Box$

We report the average $S_{\text{ES}}$ taken across all observed data.  In this form, the interpretation of $S_{\text{ES}}$ is similar to that of $S_{\text{PPL}}$: the smaller the score, the better.
% EOF