### 2022-01-05 evaluation.tex

\section{Scoring criteria for distributions in ${\mathbb S}_\infty^{d-1}$\label{sec:evaluation}}

In order to assess and compare the estimation of a distribution on ${\mathbb S}_\infty^{d-1}$ we consider the theory of proper scoring rules developed in \cite{gneiting2007}.
As mentioned in Section~\ref{subsec:projgamma}, our approach does not provide a density on ${\mathbb S}_{\infty}^{d-1}$, restricting our ability to construct model selection criteria to sample-based approaches.  We consider two scoring methods:  \emph{posterior predictive loss} (PPL) introduced in \cite{gelfand1998}, and \emph{energy score} (ES) introduced in \cite{gneiting2007}.

\subsection{Posterior predictive loss}
The posterior predictive loss criterion is introduced in \cite{gelfand1998}.  When we
  assume a squared error loss function, then for the $i$th observation, the posterior predictive
  loss criterion is computed as
  \begin{equation}
    \label{eq:ppl}
      S_{\text{PPL}}^k\left(P,x_i\right) = \text{Var}_P\left[X_i\right] + \frac{k}{k+1}\left(\text{E}_P\left[X_{i}\right] - x_i\right)^2
  \end{equation}
  where $X_i$ is a random variable from the posterior predictive distribution for $x_i$.  The
  scalar $k$ is a weighting factor that scales the importance of goodness of fit
  relative to precision.  In fact a small $\text{Var}(X_{i})$ indicates high
  precision, and a small $(\text{E}[X_{i}] - x_{i})^2$  indicates a good fit.  In our analysis,
  we take the limit as $k\to\infty$, weighting both
  equally. Overall, the smaller the score, the better.  Note that the PPLC score is defined 
  for a univariate distribution.  As we are dealing
  with a multivariate distribution, we will be taking the posterior predictive loss
  summed over all dimensions.  That is,
  \begin{equation}
    \label{eq:ppl2}
      S_{\text{PPL}}\left(P,\bm{x}_i\right) = \frac{1}{d}\sum_{\ell = 1}^d\left[\text{Var}_P\left[X_{i\ell}\right] + \left(\text{E}_P\left[X_{i\ell}\right] - x_{i\ell}\right)^2\right] .
  \end{equation}
  We report the average $S_{\text{PPL}}$ taken from all observed data.

\subsection{Energy score}
We focus on \emph{energy scores} defined for a general probability distribution $P$, with finite expectation, as 
    \begin{equation}
    \label{eq:es}
    \text{S}_{\text{ES}}\left(P, \bm{x}_i\right) =  \text{E}_p \left[g\left(\bm{X}_i, \bm{x}_i\right)\right] -
                                    \frac{1}{2}\text{E}_p \left[g\left(\bm{X}_i,\bm{X}_i^{\prime}\right)\right],
    \end{equation}
where $g$ is a kernel function. The score defined in Equation \eqref{eq:es} can be evaluated
  using samples from $P$, with the help of the law of large numbers.
  Moreover, Theorem~4 in \cite{gneiting2007}, states that if $g(\cdot,\cdot)$ is a negative 
  definite kernel, then $\text{S}(P,\bm{x})$ is a \emph{proper} scoring rule.  Recall that a 
  real valued function $g$ is a negative definite kernel if it is symmetric in its arguments, 
  and $\sum_{i=1}^n\sum_{j=1}^na_ia_jg(x_i,x_j)\leq 0$ for all positive integers $n$, and any 
  collection $a_1,\ldots,a_n\in{\mathbb R}$ such that  $\sum_{i = 1}^na_i = 0$.

In a Euclidean space, these conditions are satisfied by the Euclidean distance. However, in
  ${\mathbb S}_{\infty}^{d-1}$, Euclidean distance will tend to under-estimate the actual 
  distance required to travel between two points.  Let
\begin{equation*}
    {\mathbb C}_{\ell}^{d-1} = \lbrace \bm{x} : \bm{x} \in {\mathbb S}_{\infty}^{d-1}, x_{\ell} = 1\rbrace
\end{equation*}
comprise the $\ell$th \emph{face}.  For points on the same face, Euclidean distance 
  corresponds to the length of the shortest possible path in ${\mathbb S}_{\infty}^{d-1}$.  
  For points on different faces, the Euclidean distance is a lower bound for that length.

For a finite $p$, the shortest connecting path between two points in ${\mathbb S}_p^{d-1}$
  is the minimum geodesic; its length satisfying the definition of a distance.  Thus its
  length can be used as a negative definite kernel for the purpose of defining an energy
  score. Unfortunately as $p\to\infty$ the resulting surface ${\mathbb S}_{\infty}^{d-1}$
  is not differentiable, implying that routines to calculate geodesics are not readily
  available.  However, as ${\mathbb S}_{\infty}^{d-1}$ is a portion of of a $d$-cube, we
  can borrow a result from geometry \citep{pappas1989} stating that the length of the
  shortest path between two points on a geometric figure corresponds to the length of a
  straight line drawn between the points on an appropriate unfolding, rotation, or \emph{net} of the figure from a $d$-dimensional to a $d-1$-dimensional space.  The optimal net will have the shortest straight line between the points, as long as that line is fully contained within such net. As ${\mathbb S}_{\infty}^{d-1}$ has $d$ faces---each face pairwise adjacent, there are $d!$ possible nets.  However, we are only interested in nets that begin and end on the source and destination faces respectively.  That reduces the number of nets under consideration to $\sum_{k = 0}^{d-2}\binom{d-2}{k}$.  This is still computationally burdensome for a large number of dimensions.

To reduce the computations needed to calculate the energy score we define a kernel as 
in the following proposition.
\begin{prop}\label{prop:g}
Let $\bm{a} \in {\mathbb C}_{\ell}^{d-1}$, and $\bm{b} \in {\mathbb C}_{\jmath}^{d-1}$, for $\ell, \jmath \in \{1, \ldots , d\}$. Define $g$ as
\[  
    g(\bm{a},\bm{b}) = \min_{\bm{c} \in{\mathbb C}_{\jmath}^{d-1}\cap{\mathbb C}_{\ell}^{d-1}}\left\{ 
        \pnorm{\bm{c}-\bm{a}}{2} + \pnorm{\bm{b}-\bm{c}}{2} \right\}.
\]
Then $g$  is a negative definite kernel.
\end{prop}

{\em Proof:}
When $\bm{a}$ and $\bm{b}$ are on the same face, then $\ell = \jmath$, and $g$ is the Euclidean distance between $\bm{a}$ and $\bm{b}$, which is a  negative definite kernel.  When $\bm{a}$ and $\bm{b}$ are on separate
  faces, let 
  \[\bm{e} = 
    \argmin_{\bm{c} \in{\mathbb C}_{\jmath}^{d-1}\cap{\mathbb C}_{\ell}^{d-1}}\left\{ 
        \pnorm{\bm{c}-\bm{a}}{2} + \pnorm{\bm{b}-\bm{c}}{2} \right\}.
  \]
  We have that $g(\bm{a}, \bm{b}) = g(\bm{b}, \bm{a})$ due to the uniqueness of $\bm{e}$ and the symmetry of the Euclidean distance. Additionally, given $n$ and  a collection $\alpha_1,\ldots,\alpha_n\in{\mathbb R}$ such that  $\sum_{i = 1}^n\alpha_i = 0$, then, 
  \begin{equation*}
    \begin{aligned}
      \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j g(\bm{x}_1,\bm{x}_2) &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j \bigg[\pnorm{\bm{e} - \bm{x}_1}{2} + \pnorm{\bm{x}_2 - \bm{e}}{2}\bigg]\\
      &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{e} - \bm{x}_1}{2} + \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{x}_2 - \bm{e}}{2} \leq0,
    \end{aligned}
  \end{equation*}
  as the Euclidean norm is a negative definite kernel.$\hfill\Box$

Evaluating $g$ as defined in Proposition \ref{prop:g} requires the solution of a $(d-2)$-dimensional 
  optimization problem.  The following proposition provides a straightforward solution.
\begin{prop}\label{prop:rot}
    Let $\bm{a} \in {\mathbb C}_{\ell}^{d-1}$, and $\bm{b} \in {\mathbb C}_{\jmath}^{d-1}$, for $\ell, \jmath \in \{1, \ldots , d\}$.  For $\ell\neq \jmath$, the 
  transformation $P_{\jmath\ell}(\cdot)$ required to rotate the $\jmath$th face along the $\ell$th axis 
  produces the vector $\bm{b}^\prime$, where
  \begin{equation}
    \label{eqn:rotation}
     \bm{b}^{\prime}_i = P_{\jmath\ell}(\bm{b}) = 
    \begin{cases}
        b_{i} &\text{for }i\neq \jmath,\ell\\
        1 &\text{for }i = \ell\\
        2 - b_{\ell} &\text{for }i = \jmath
    \end{cases}.
  \end{equation}
  Then $g(\bm{a},\bm{b}) = \pnorm{\bm{a} - \bm{b}^{\prime}}{2}$.
\end{prop}
{\em Proof:}
Notice that for 
  $\bm{c} \in {\mathbb C}_{\jmath}^{d-1}\cap{\mathbb C}_{\ell}^{d-1}$,
     $\pnorm{\bm{b} - \bm{c}}{2} =  \pnorm{\bm{b}^{\prime} - \bm{c}}{2}$. Then, recalling the definition of 
     $\bm{e}$ in the proof of Proposition \ref{prop:g}, we have that
  \begin{equation*}
    \begin{aligned}
    g(\bm{a}, \bm{b}) &= \pnorm{\bm{a} - \bm{e}}{2} + \pnorm{\bm{b} - \bm{e}}{2}\\
    &= \pnorm{\bm{a} - \bm{e}}{2} + \pnorm{\bm{b}^{\prime} - \bm{e}}{2} = \pnorm{\bm{a} - \bm{b}^{\prime}}{2}.
    \end{aligned}
  \end{equation*}
  The last equality is due to the fact that $\bm{a}$ and $\bm{b}'$ belong to the same hyperplane.
  $\hfill\Box$

Proposition \ref{prop:rot} provides a computational efficient way to evaluate the proper scoring rule $S_{\text{ES}}$ defined on ${\mathbb S}_\infty^{d-1}$, for each observation. For the purpose of model assessment and comparison, we report the average $S_{\text{ES}}$ taken across all observed data.  In this form, the interpretation of $S_{\text{ES}}$ is similar to that of $S_{\text{PPL}}$: The smaller the score, the better.
% EOF