### 2022-01-05 main.tex

\documentclass[12pt,authoryear]{elsarticle}
%\documentclass[12pt]{article}
% \usepackage[margin=1in]{geometry}
% \usepackage[margin=0.5in]{caption}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{natbib}
\usepackage{amsthm}
\usepackage{hyperref}
%\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{soul}
\usepackage{algorithm2e}
\usepackage{bm}
\usepackage{subcaption}
\usepackage{wrapfig}
% \usepackage{setspace}
% \linespread{1.8}
\usepackage{anysize}
\marginsize{1in}{1in}{1in}{1in}

\newtheorem{prop}{Proposition}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\bruno}[1]{\textcolor{blue}{#1}} % Bruno Sanso edits - blue
\newcommand{\makenote}[1]{\textcolor{red}{#1}} % Peter edits - red
\newcommand{\findcite}{\textcolor{red}{[Find Citation]}}
\newcommand{\needcite}{\findcite}
\newcommand{\Chi}{\mbox{\Large$\chi$}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\inorm}[1]{\norm{#1}_{\infty}}
\newcommand{\pnorm}[2]{\norm{#1}_{#2}}
\newcommand{\prob}[1]{\text{Pr}\left[#1\right]}
\newcommand{\expect}[1]{\text{E}\left[#1\right]}

\title{Inference for multivariate peaks over threshold models}
\author{Peter Trubey}
\ead{ptrubey@ucsc.edu}
%\address{UC Santa Cruz, 1156 High St., Santa Cruz, CA 95064, USA}

\author{Bruno Sans\'o}
\ead{bruno@soe.ucsc.edu}
\address{Department of Statistics, University of California, Santa Cruz,  USA}

\begin{document}

%--------title page
% \thispagestyle{empty}
\begin{abstract}
  \input{abstract}
\end{abstract}

\begin{keyword}
 Multivariate extremes \sep Peak over threshold models \sep
 Bayesian non-parametric models \sep Dirichlet process mixtures
\end{keyword}

\maketitle

%--------content
\input{introduction}

\input{methodology}

\input{evaluation}

\input{results}

\input{conclusion}

\bibliographystyle{elsarticle-harv}
\bibliography{./refs}

\end{document}

% EOF