### 2022-05-27 cover.tex

\documentclass[letterpaper,11pt]{letter}

\name{Peter Trubey}
\signature{Peter Trubey}
\address{Peter Trubey \\ Department of Statistics \\ University of California, Santa Cruz \\ USA}

\begin{document}%

\begin{letter}{Editorial Board \\ Computational Statistics \& Data Analysis}

\opening{Dear Editorial Board,}

We submit to you our manuscript for review.  It details a method of inference for the dependence structure of
extremes in a multivariate peaks over thresholds model based mixtures of projections of vectors of gamma variables, 
along with an appropriate model selection criteria and applications of the fitted model.  A simulation study 
is performed to explore properties of the distribution, and the model then applied to atmospheric data.

We confirm the content of this work is original, and is not under consideration at any other journal.

We further confirm that there are no competing interests to declare.

Sincerely,

\closing{~}

\end{letter}
\end{document}

% EOF