### 2022-05-27 main.tex

\documentclass[12pt,authoryear]{elsarticle}
%\documentclass[12pt]{article}
% \usepackage[margin=1in]{geometry}
% \usepackage[margin=0.5in]{caption}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{natbib}
\usepackage{amsthm}
\usepackage{hyperref}
%\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{soul}
\usepackage[ruled]{algorithm2e}
\usepackage{bm}
\usepackage{subcaption}
\usepackage{wrapfig}
% \usepackage{setspace}
% \linespread{1.8}
\usepackage{anysize}
\marginsize{1in}{1in}{1in}{1in}

\newtheorem{prop}{Proposition}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\bigmid}{\,\middle\vert\,}
% \newcommand{\bruno}[1]{\textcolor{blue}{#1}} % Bruno Sanso edits - blue
% \newcommand{\makenote}[1]{\textcolor{red}{#1}} % Peter edits - red

\newcommand{\findcite}{\textcolor{red}{[Find Citation]}}
\newcommand{\needcite}{\findcite}
\newcommand{\Chi}{\mbox{\Large$\chi$}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\inorm}[1]{\norm{#1}_{\infty}}
\newcommand{\pnorm}[2]{\norm{#1}_{#2}}
\newcommand{\prob}[1]{\text{Pr}\left[#1\right]}
\newcommand{\expect}[1]{\text{E}\left[#1\right]}

\SetAlCapNameFnt{\footnotesize}
\SetAlCapFnt{\footnotesize}
\SetAlCapSty{}

\begin{document}

\begin{frontmatter}
    %\title{Inference for multivariate peaks over threshold models}
    \title{Bayesian non-parametric inference for multivariate peaks-over-threshold models}
    \author{Peter Trubey\corref{cor1}}
    \ead{ptrubey@ucsc.edu}

    \author{Bruno Sans\'o}
    \ead{bruno@soe.ucsc.edu}
    \address{Department of Statistics, University of California, Santa Cruz,  USA}
    \cortext[cor1]{Corresponding author}

    \begin{abstract}
        \input{abstract}
    \end{abstract}

    \begin{keyword}
        Multivariate extremes \sep Peak over threshold models \sep
        Bayesian non-parametric models \sep Dirichlet process mixtures
    \end{keyword}
\end{frontmatter}

%--------content
\input{introduction}

\input{methodology}

\input{evaluation}

\input{results}

\input{conclusion}

\section*{Acknowledgements}
This material is based upon work supported by the National Science Foundation under Grant No. SES-2050012 and DMS-2153277.

\bibliographystyle{elsarticle-harv}
\bibliography{./refs}

\end{document}

% EOF