### 2022-06-15 abstract.tex
first pass at springer-nature class

% We consider a constructive definition of the multivariate Pareto that factorizes the 
% random vector into a radial component and an independent angular component; The former
% following a univariate Pareto distribution, and the latter defined on the surface of the 
% positive orthant of the unit hypercube.  In this paper, we  propose a method for
% inferring the distribution of the angular component.  We identify its support 
% as the limit of the positive orthants of the unit $p$--norm spheres,
% and introduce a projected gamma family of
% distributions defined as the projection of a vector of independent gamma random variables
% onto the $p$--norm sphere.  This family serves as a building block for a flexible family
% of distributions obtained as a Dirichlet process mixture of projected gammas.  For 
% model assessment and comparison, we discuss model scoring methods appropriate 
% to distributions on the unit hypercube.  In particular, working with the energy score
% criterion, we develop a kernel metric appropriate to the hypercube that produces a 
% proper scoring rule.  We then present a simulation study to compare 
% different modeling choices using the proposed scoring rules.  Finally we apply our 
% approach to describe the dependence structure of the extreme values of the magnitude 
% of the integrated vapor  transport (IVT), a variable that describes the rate of flow 
% of moisture in the atmosphere  along the coast of California for the years of 1979 
% through 2020.  We find a clear but heterogeneous geographical dependence.

We consider a constructive definition of the multivariate Pareto that factorizes the 
random vector into a radial component and an independent angular component; The former
following a univariate Pareto distribution, and the latter defined on the surface of the 
positive orthant of the unit hypercube.  In this paper, we  propose a method for
inferring the distribution of the angular component.  We identify its support 
as the limit of the positive orthant of the unit $p$--norm spheres,
and introduce a projected gamma family of
distributions defined as the projection of a vector of independent gamma random variables
onto the $p$--norm sphere.  This family serves as a building block for a flexible family
of distributions obtained as a Dirichlet process mixture of projected gammas.  For 
model assessment, we discuss scoring methods appropriate 
to distributions on the unit hypercube.  In particular, working with the energy score
criterion, we develop a kernel metric that produces a 
proper scoring rule, and present a simulation study to compare 
different modeling choices using the proposed scoring rules.  Finally we apply our 
approach to describe the dependence structure of the extreme values of the magnitude 
of the integrated vapor transport (IVT), data describing the flow 
of atmospheric moisture along the coast of California for the years of 1979 
through 2020.  We find clear but heterogeneous geographical dependence.

% EOF