### 2022-06-15 main.tex

\documentclass[sn-basic]{sn-jnl}
%\documentclass[12pt,authoryear]{elsarticle}
%\documentclass[12pt]{article}
% \usepackage[margin=1in]{geometry}
% \usepackage[margin=0.5in]{caption}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{natbib}
\usepackage{amsthm}
\usepackage{hyperref}
%\usepackage{graphicx}
%\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{soul}
\usepackage[ruled]{algorithm2e}
\usepackage{bm}
\usepackage{subcaption}
\usepackage{wrapfig}
% \usepackage{setspace}
% \linespread{1.8}
\usepackage{anysize}
\marginsize{1in}{1in}{1in}{1in}

\newtheorem{prop}{Proposition}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\bigmid}{\,\middle\vert\,}
% \newcommand{\bruno}[1]{\textcolor{blue}{#1}} % Bruno Sanso edits - blue
% \newcommand{\makenote}[1]{\textcolor{red}{#1}} % Peter edits - red

\newcommand{\findcite}{\textcolor{red}{[Find Citation]}}
\newcommand{\needcite}{\findcite}
\newcommand{\Chi}{\mbox{\Large$\chi$}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\inorm}[1]{\norm{#1}_{\infty}}
\newcommand{\pnorm}[2]{\norm{#1}_{#2}}
\newcommand{\prob}[1]{\text{Pr}\left[#1\right]}
\newcommand{\expect}[1]{\text{E}\left[#1\right]}

% \SetAlCapNameFnt{\footnotesize}
% \SetAlCapFnt{\footnotesize}
% \SetAlCapSty{}

\jyear{2022}
\theoremstyle{thmstyleone}

\raggedbottom

\begin{document}

\title[Bayesian non-parametric inference for multivariate PoT models]{Bayesian non-parametric inference for multivariate peaks-over-threshold models}

\author*[1]{\fnm{Peter} \sur{Trubey}}\email{ptrubey@ucsc.edu}
\author[1]{\fnm{Bruno} \sur{Sans\'{o}}}\email{bruno@soe.ucsc.edu}

\affil*[1]{\orgdiv{Department of Statistics}, \orgname{University of California Santa Cruz}, 
            \orgaddress{\street{1156 High Street}, \city{Santa Cruz},\postcode{95064}, \state{CA}, \country{USA}}}

\abstract{\input{./abstract}}
\keywords{Multivariate extremes, Peak over threshold models,
                Bayesian non-parametric models, Dirichlet process mixtures}

% \Author{Peter Trubey\Affil{1} and Bruno Sans\'o\Affil{1}}
% \AuthorRunning{Peter Trubey and Bruno Sans\'o}

% \Affiliations{
% \item Department of Statistics,
%       University Of California Santa Cruz,
%       USA
% }

% \CorrAddress{Peter Trubey,
%             Department of Statistics,
%             University of California Santa Cruz,
%             1156 High Street, 
%             Santa Cruz, CA 95064,
%             USA}
% \CorrEmail{ptrubey@ucsc.edu}
% \CorrPhone{831-459-1448}
% \CorrFax{831-459-4829}

% \Title{Bayesian non-parametric inference for multivariate peaks-over-threshold models}
% \TitleRunning{Bayesian non-parametric inference for multivariate PoT models}

% \Abstract{
%     \input{./abstract}
% }

% \Keywords{
%  Multivariate extremes; Peak over threshold models; 
%  Bayesian non-parametric models; Dirichlet process mixtures
% }

\maketitle
https://www.overleaf.com/project/612e69c665059f5fd7998d90
% \begin{frontmatter}
%     %\title{Inference for multivariate peaks over threshold models}
%     \title{Bayesian non-parametric inference for multivariate peaks-over-threshold models}
%     \author{Peter Trubey\corref{cor1}}
%     \ead{ptrubey@ucsc.edu}

%     \author{Bruno Sans\'o}
%     \ead{bruno@soe.ucsc.edu}
%     \address{Department of Statistics, University of California, Santa Cruz,  USA}
%     \cortext[cor1]{Corresponding author}

%     \begin{abstract}
%         \input{abstract}
%     \end{abstract}

%     \begin{keyword}
%         Multivariate extremes \sep Peak over threshold models \sep
%         Bayesian non-parametric models \sep Dirichlet process mixtures
%     \end{keyword}
% \end{frontmatter}

%--------content
\input{./introduction}

\input{./methodology}

\input{./evaluation}

\input{./results}

\input{./conclusion}

\section*{Acknowledgements}
This material is based upon work supported by the National Science Foundation under Grant No. SES-2050012 and DMS-2153277.

%\bibliographystyle{elsarticle-harv}
\bibliography{./refs}

\end{document}

% EOF