### 2022-06-18 conclusion.tex

\section{Conclusion\label{sec:conclusion}}
In this paper, we have built upon the definition of the multivariate Pareto described in \cite{ferreira2014}
  to establish a useful representation of its dependence structure, through
  the distribution of its angular component, which is supported on the positive orthant of the unit
  hypersphere under the $\mathcal{L}_{\infty}$ norm.  Due to the 
  inherent difficulty of obtaining a distribution with support on ${\mathbb S}^{d-1}_\infty$
  our method projects distributions, supported on the positive quadrant and
  based on products of independent gammas,  onto the manifold
  ${\mathbb S}_{p}^{d-1}$. Samples of the resulting probability distribution 
  are then projected onto ${\mathbb S}_{\infty}^{d-1}$.  
  As ${\mathbb S}_{p}^{d-1}$ converges to ${\mathbb S}_{\infty}^{d-1}$ as $
  p\to\infty$, we expect the resampling to be efficient for large enough
  $p$. In fact, our exploration of the simulated and real data indicate
  that the procedure is robust to the choice of moderately large values
  of $p$.

We explored the behavior of the proposed model using simulated data from
    mixtures of projected Gammas with varying degrees of complexity---both in terms of dimensionality and 
    in number of mixture components.  From this, we learned that the additional flexibility 
    offered by varying the $\beta$ term---the rate parameter of the Gamma distribution---does not result 
    in increased model fidelity in a mixture setting.  Within the tested range, between 3 and 20 
    dimensions, we observed that the additional information provided by a log--normal 
    centering distribution did not translate to additional model fidelity.  However, for 
    real data, with 47 dimensions, we did observe an effect.

The computations in this paper were performed on a desktop computer with an AMD Ryzen 5950X processor.
  The program is largely single-threaded, so computation time is not dependent on available core count.  In each
  case, we run the MCMC chain for 50\,000 iterations, with a burn-in of 20\,000 samples.  Fitting the PG--G model
  on the ERA5 dataset took approximately 15 minutes.  Work is in progress to optimize the code, and explore
  parallelization where possible.  We are also exploring alternative computational approaches that will make it
  feasible to tackle very high dimensional problems. In fact, to elaborate on the study of IVT, there is a need 
  to consider several hundreds, if not thousands, of grid cells over the Pacific Ocean in order to obtain a 
  good description of atmospheric events responsible for large storm activity over California.