### 2022-06-18 main.tex

\documentclass[iicol,sn-basic]{sn-jnl}

\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{natbib}
\usepackage{amsthm}
\usepackage{hyperref}
%\usepackage{graphicx}
%\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{soul}
%\usepackage[ruled]{algorithm2e}
\usepackage{bm}
\usepackage{subcaption}
\usepackage{wrapfig}
% \usepackage{setspace}
% \linespread{1.8}
\usepackage{anysize}
\marginsize{1in}{1in}{1in}{1in}

\newtheorem{prop}{Proposition}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\bigmid}{\,\middle\vert\,}
% \newcommand{\bruno}[1]{\textcolor{blue}{#1}} % Bruno edits - blue
% \newcommand{\makenote}[1]{\textcolor{red}{#1}} % Peter edits - red

\newcommand{\findcite}{\textcolor{red}{[Find Citation]}}
\newcommand{\needcite}{\findcite}
\newcommand{\Chi}{\mbox{\Large$\chi$}}

\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\inorm}[1]{\norm{#1}_{\infty}}
\newcommand{\pnorm}[2]{\norm{#1}_{#2}}
\newcommand{\prob}[1]{\text{Pr}\left[#1\right]}
\newcommand{\expect}[1]{\text{E}\left[#1\right]}

\jyear{2022}
\theoremstyle{thmstyleone}

\raggedbottom

\begin{document}

\title[~\hspace{6.75in}BNP inference for multivariate PoT]{Bayesian non-parametric inference for multivariate peaks-over-threshold models}

\author*{\fnm{Peter} \sur{Trubey}}\email{ptrubey@ucsc.edu}
\author{\fnm{Bruno} \sur{Sans\'{o}}}\email{bruno@soe.ucsc.edu}

\affil{\orgdiv{Department of Statistics}, \orgname{University of California Santa Cruz}, 
            \orgaddress{\street{1156 High Street}, \city{Santa Cruz},\postcode{95064}, \state{CA}, \country{USA}}}

\abstract{\input{./abstract}}
\keywords{Multivariate extremes, Peak over threshold models,
                Bayesian non-parametric models, Dirichlet process mixtures}

\maketitle

%--------content
\input{./introduction}

\input{./methodology}

\input{./evaluation}

\input{./results}

\input{./conclusion}

\section*{Acknowledgements}
This material is based upon work supported by the National Science Foundation
    under Grant No. SES-2050012 and DMS-2153277.

\bibliography{./refs}

\end{document}

% EOF