### 2022-06-18 methodology_projgamma.tex

\subsection{Projected gamma family\label{subsec:projgamma}}
  A natural  distribution to consider in ${\mathbb R}^d_+$ is given by a product of independent
  univariate Gamma distributions. Let
    $\bm{ X} \sim \prod_{\ell = 1}^d\text{Ga}\left(X_{\ell}\mid\alpha_{\ell},\beta_{\ell}\right)$, 
    $\alpha_\ell$ and $\beta_\ell$ are the shape and scale parameters, respectively. Using Equations (\ref{eqn:pnormt}) -- (\ref{eqn:pnormjac}) we have the joint density
  \begin{equation}
  \label{joint}
    \begin{aligned}
    f(r,\bm{ y}) = &\prod_{\ell = 1}^{d}
      \left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}(ry_{\ell})^{\alpha_{\ell} - 1}
          \exp\lbrace-\beta_{\ell}ry_{\ell}\rbrace\right] \\
      &\hspace{0.5cm}\times r^{d-1}\left[y_d +
            {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right].
    \end{aligned}
  \end{equation}
  Integrating out $r$ yields the resulting \emph{Projected Gamma} density
  \begin{equation}
    \label{eqn:projgamma}
    \begin{aligned}
    \text{PG}(\bm{ y}\mid\bm{ \alpha},\bm{ \beta}) &=
          \prod_{\ell = 1}^d\left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}
                y_{\ell}^{\alpha_{\ell} - 1}\right]\\
      &\hspace{0.5cm}\times \left[y_d +
          {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right]\\
      &\hspace{0.5cm}\times \frac{\Gamma({\textstyle\sum}_{\ell = 1}^d\alpha_{\ell})}{\left({\textstyle\sum}_{\ell = 1}^d
                    \beta_{\ell}y_{\ell}\right)^{{\scriptstyle\sum_{\ell = 1}^d \alpha_{\ell}}}},
    \end{aligned}
  \end{equation}
  defined for $\bm{y}\in {\mathbb S}_p^{d-1}$, and for any $p>0$. To avoid identifiability problems when estimating the shape and scale parameters, we set $\beta_1 = 1$.
  \cite{nunez2019} obtain the density in Equation (\ref{eqn:projgamma})
  for $p=2$ as a multivariate distribution for directional data, using spherical coordinates.  
  For $\bm{ y}\in {\mathbb S}_1^{d-1}$, and $\beta_{\ell} = \beta$ for all $\ell$, the density in Equation (\ref{eqn:projgamma}) corresponds to that of a Dirichlet distribution.

  The projected gamma family is simple to specify and has very tractable computational properties. Thus, we use it as a building block for the angular measure $\Phi$ models. To build a flexible family of distributions in ${\mathbb S}_p^{d-1}$ we consider mixtures of projected gamma densities defined as
  \begin{equation} \label{eqn:PGmix}     
     f(\bm{y}) = \int_\Theta \mathcal{PG}(\bm{y}\mid \bm{\theta}) dG(\bm{\theta}), 
  \end{equation} 
  where $\bm{\theta} = (\bm{\alpha}, \bm{\beta})$. Following a Bayesian non-parametric approach \citep{Ferguson74,Antoniak1974} we assume that $G$ is drawn from a random measure. In particular, assuming a Dirichlet process prior for $G$, we have a hierarchical formulation of the mixture model that, for a vector of observations $\bm{y}_i$, is given by
  \begin{equation}\label{eqn:dppg}
    \begin{aligned}
    \bm{y}_i &\sim \text{PG}(\bm{y}_i\mid \bm{\theta}_i)\\
    \bm{\theta}_i &\sim G
    \end{aligned}
    ~\hspace{1cm}
    \begin{aligned}
    G &\sim \mathcal{DP}(\eta, G_0),
    \end{aligned}
  \end{equation}
  where $\mathcal{DP}$ denotes a Dirichlet process, $\eta$ is the precision parameter, and $G_0$ is the centering distribution.

  Unfortunately, in the limit when $p\rightarrow \infty$, the 
  projection transformation is not differentiable. Thus, a closed form expression 
  like Equation (\ref{eqn:projgamma}) for the projected gamma density on ${\mathbb S}_\infty^{d-1}$ is not 
  available. Instead, we observe that for a sufficiently large $p$, $\mathbb{S}_p^{d-1}$ will approach
  $\mathbb{S}_{\infty}^{d-1}$.  With that in mind, our strategy consists of describing the angular
  distribution $\Phi$ using a sample based approach with the following steps: (i) Apply the 
  transformation in Equation (\ref{eqn:standardization}) to the original data; (ii) Obtain the 
  subsample of the standardized observations that satisfy $R>1$; (iii) Take a finite $p$ and 
  project the observations onto ${\mathbb S}_p^{d-1}$; 
  (iv) Fit the model in Equation (\ref{eqn:PGmix}) to the resulting 
  data and obtain samples from the fitted model; 
  (v) Project the resulting samples onto ${\mathbb S}_\infty^{d-1}$.
  For step (iv) we use a Bayesian approach that is implemented using a purposely developed Markov chain Monte Carlo sampler described in the next section.

\begin{figure}[ht]
    \centering
    \caption{The positive orthant of the $p$-norm sphere.}
  \begin{subfigure}[b]{0.4\textwidth}
    \includegraphics[width=\textwidth]{./images/p_sphere}
    \caption{${\mathbb S}_p^1$ for $p = 1,\ldots, 10$\label{fig:psphere}}
  \end{subfigure}
  %
  \begin{subfigure}[b]{0.4\textwidth}
    \includegraphics[width=\textwidth]{./images/p_project}
    \caption{Projection of data on ${\mathbb S}_{\infty}^1$ onto ${\mathbb S}_p^1$\label{fig:pproject}}
  \end{subfigure}
\end{figure}

\subsubsection{Inference for the projected Gamma mixture model}
In a sample-based inference approach, for a given iteration the Dirichlet process mixture model groups observations
    into stochastically assigned clusters, where members of a cluster share distributional parameters. Building out 
    the methods of inference for Equation~(\ref{eqn:dppg}), 
    let $n_j^{-i}$ be the number of observations in cluster $j$ not including observation $i$.  Let $J^{-i}$ be 
    the number of extant clusters, not including any singleton that observation $i$ is in. Under this model, 
    the probability of cluster membership for a given observation is proportional to
\begin{equation*}
    \text{Pr}\left[\delta_i = j\mid\ldots\right] \propto \begin{cases}
        n_j^{-i}\mathcal{PG}\left(\bm{y}_i\mid\bm{\alpha}_j,\bm{\beta}_j\right)\\ % \hspace{0.5cm} \\ % &\text{~for~}j \in \lbrace 1,\ldots,J^{-i}\rbrace\\
        \eta\int\mathcal{PG}\left(\bm{y}_i\mid\bm{\alpha}_j,\bm{\beta}_j\right)dG_0(\bm{\alpha}_j,\bm{\beta}_j),%\hspace{0.5cm} % &\text{~for~}j = J^{-i} + 1.
        \end{cases}
\end{equation*}
where the top case is iterating over extant clusters $j = 1,\ldots, J^{-i}$, and the bottom case is
    for a \emph{new} cluster. If $G_0$ is not a conjugate prior for the kernel density, the integral in the 
    above formula may not be available in closed form. We sidestep this using Algorithm 8 from \cite{neal2000}: 
    by Monte Carlo integration, we draw $m$ candidate clusters, $\bm{\alpha}_j,\bm{\beta}_j$ for
    $j = J^{-i} + 1,\ldots, J^{-i} + m$ from $G_0$. Then, we sample the cluster indicator 
    $\gamma_i$ from extant or candidate clusters, where the probability of cluster membership is proportional to
\begin{equation}
    \text{Pr}\left[\delta_i = j\mid\ldots\right] \propto \begin{cases}
        n_j^{-i}\mathcal{PG}\left(\bm{y}_i\mid\bm{\alpha}_j,\bm{\beta}_j\right)\\ % \hspace{0.5cm}  % &\text{~for~}j \in \lbrace 1,\ldots,J^{-i}\rbrace\\
        \frac{\eta}{m}\mathcal{PG}\left(\bm{y}_i\mid\bm{\alpha}_j,\bm{\beta}_j\right). % \hspace{0.5cm} &\text{~for~}j \in \lbrace J^{-i} + 1,\ldots, J^{-i} + m\rbrace.
        \end{cases}
\end{equation}
Again, the top case is iterating over extant clusters, and now the bottom case is iterating 
    over new \emph{candidate} clusters.  If a candidate cluster is selected,
%{\bf you first have to say that you sample from this distribution. What is the ``candidate cluster''?}
    then $\gamma_i = J = J^{- i} + 1$, and the associated cluster parameters are saved.

A key feature of the the projected Gamma distribution is its computational properties. We augment 
$\mathcal{PG}(\bm{y}_i\mid\bm{\alpha}_i,\bm{\beta}_i) $ by introducing a latent radial component $r_i$, 
for each observation. Using Equation \eqref{joint} we observe that the 
full conditional of $r_i$ is easy to sample from, as it is given as
\begin{equation}
    r_i\mid\bm{\alpha}_i,\bm{\beta}_i \sim \mathcal{G}\left(r_i \bigmid \sum_{\ell = 1}^d\alpha_{i\ell}, \sum_{\ell = 1}^d\beta_{\ell} y_{i\ell}\right).
\end{equation}
Moreover,  the full conditional for $\bm{\alpha}_j,\bm{\beta}_j$ is then proportional to
\begin{equation}
    \begin{aligned}
    &\hspace{-1cm}f(\bm{\alpha}_j,\bm{\beta}_j\mid \bm{Y},\bm{r},\bm{\delta},\ldots) \\
    &\propto \prod_{i:\gamma_i = j}\prod_{\ell = 1}^d\mathcal{G}\left(r_iy_{i\ell}\mid\alpha_{j\ell},\beta_{j\ell}\right) \\
    &\hspace{0.5cm} \times dG_0(\bm{\alpha}_j,\bm{\beta}_j).
    \end{aligned}
\end{equation}
Note that the ordering of the products can be reversed in the likelihood, indicating that given $\bm{r}$, 
the likelihood becomes separable by dimension.
We first consider a centering distribution given by a product of independent Gammas:
\begin{equation}
    \begin{aligned}
    &\hspace{-0.3cm}G_0(\bm{\alpha}_j,\bm{\beta}_j\mid \bm{\xi},\bm{\tau},\bm{\zeta},\bm{\sigma}) \\
    &= \prod_{\ell = 1}^d\mathcal{G}(\alpha_{j\ell}\mid \xi_{\ell},\tau_{\ell})\times\prod_{\ell = 2}^d\mathcal{G}(\beta_{j\ell}\mid\zeta_{\ell},\sigma_{\ell}).
    \end{aligned}
\end{equation}
This model is completed with independent Gamma priors on $\xi_{\ell}$, $\tau_{\ell}$, $\zeta_{\ell}$, $\sigma_{\ell}$.  We also assume a Gamma prior on $\eta$, that is updated via the procedure outlined in \cite{escobar1995}.  We refer to this model as the \emph{projected Gamma--Gamma} (PG--G) model.  An advantage of the PG--G model is that, thanks to conjugacy, the rate parameters can easily be integrated out for inference on $\bm{\alpha}_j$.  Then, the full conditional for $\alpha_{j\ell}$ takes the form
\begin{equation}
    \label{eqn:alphalupdate}
    \begin{aligned}
    &\hspace{-0.5cm}\pi(\alpha_{j\ell}\mid \bm{r},\bm{Y},\bm{\gamma},\xi_\ell,\tau_\ell,\zeta_\ell,\sigma_\ell) \\
    &\propto \frac{\left(\prod_{i:\gamma_i = j}r_iy_{i\ell}\right)^{\alpha_{j\ell} - 1}\alpha_{j\ell}^{\xi_\ell - 1}e^{-\tau_\ell \alpha_{j\ell}}\times}{\Gamma^{n_j}(\alpha_{j\ell})} \\
    &\hspace{0.5cm}\times \frac{\Gamma\left(n_j\alpha_{j\ell} + \zeta_{\ell}\right)}{\left(\sum_{i:\gamma_i = j}r_iy_{i\ell} + \sigma_{\ell}\right)^{n_j\alpha_{j\ell} + \zeta_{\ell}}}
    \end{aligned}
\end{equation}
for $\ell = 2,\ldots,d$.  For $\ell = 1$, as $\beta_{1} := 1$, the full conditional takes the simpler form
\begin{equation}
    \label{eqn:alpha1update}
    \begin{aligned}
    &\hspace{-0.5cm}\pi(\alpha_{j1}\mid\bm{r},\bm{Y},\bm{\gamma},\xi_1,\tau_1) \\
    &\propto \frac{\left(\prod_{i:\gamma_i = j}r_iy_{i1}\right)^{\alpha_{j1} - 1}\alpha_{j1}^{\xi_1 - 1}e^{-\tau_1\alpha_{j1}}}{\Gamma^{n_j}(\alpha_{j1})}.
    \end{aligned}
\end{equation}
Samples of $\alpha_{j\ell}$ can thus be obtained using a Metropolis step. In our analysis, we first transform $\alpha_{j\ell}$ to the log scale, and use a normal proposal density.
The full conditional for $\beta$ is 
\begin{equation}
    \label{eqn:betafc}
    \begin{aligned}
    &\hspace{-0.5cm}\beta_{j\ell}\mid\bm{r},\bm{Y},\alpha,\zeta_{\ell},\sigma_{\ell} \\
    &\sim \mathcal{G}\left(\beta_{j\ell}\bigmid n_j\alpha_{j\ell} + \zeta_\ell, \sum_{i:\gamma_i = j}r_iy_{i\ell} + \sigma_{\ell}\right),
    \end{aligned}
\end{equation}
for $\ell = 2,\ldots, d$.  Updating $\beta_{j\ell}$ is done via a Gibbs step.  The hyper-parameters $\xi_{\ell},\tau_{\ell},\zeta_{\ell},\sigma_{\ell}$ follow similar Gamma-Gamma update relationships.  We also explore a restricted form of this model, where $\beta_{\ell} = 1$ for all $\ell$.  Under this model, we use the full conditional in Equation~(\ref{eqn:alpha1update}) for all $\ell$, and omit inference on $\bm{\zeta},\bm{\sigma}$.  We refer to this model as the \emph{projected restricted Gamma--Gamma} (PRG--G) model.

The second form of centering distribution we explore is a multivariate log-normal distribution on the shape parameters $\bm{\alpha}_j$, with independent Gamma $\beta_{j\ell}$ rate parameters.  
\begin{equation}
    \begin{aligned}
    &\hspace{-0.5cm}G_0\left(\bm{\alpha}_j,\bm{\beta}_j\mid\bm{\mu},\bm{\Sigma},\zeta,\sigma\right) \\
    &=\mathcal{LN}\left(\bm{\alpha}_j\mid\bm{\mu},\bm{\Sigma}\right)\times\prod_{\ell = 2}^d\mathcal{G}\left(\beta_{j\ell}\mid\zeta_{\ell},\sigma_{\ell}\right).
    \end{aligned}
\end{equation}
This model is completed with a normal prior on $\bm{\mu}$, an inverse Wishart prior on $\bm{\Sigma}$, and Gamma priors on $\zeta_{\ell}$ and $\sigma_{\ell}$, and $\eta$.  This model is denoted as the \emph{projected Gamma--log-normal} (PG--LN) model.  We also explore a restricted Gamma form of this model as above, where $\beta_{\ell} = 1$ for all $\ell$.  This is denoted as the \emph{projected restricted Gamma--log-normal} (PRG--LN) model.  Updates for $\bm{\alpha}$ can be accomplished using a joint Metropolis step, where $\beta_{j\ell}$ for $\ell = 2,\ldots,d$ have been integrated out of the log-density.  That is,
\begin{equation*}
    \begin{aligned}
    &\hspace{-0.5cm}\pi(\bm{\alpha}_j\mid\bm{Y},\bm{r},\bm{\delta},\bm{\mu},\bm{\Sigma},\bm{\zeta},\bm{\sigma})\\
    &\propto \exp\left\lbrace-\frac{1}{2}(\log\bm{\alpha}_j - \mu)^T\bm{\Sigma}^{-1}(\log\bm{\alpha}_j - \mu)\right\rbrace \\
    &\hspace{0.5cm}\times\frac{1}{\prod_{\ell = 1}^d\alpha_{j\ell}}\times \frac{\left(\prod_{i:\gamma_i = j}r_iy_{i1}\right)^{\alpha_{j1} - 1}}{\prod_{\ell = 1}^d\Gamma^{n_j}(\alpha_{j\ell})}\\
    &\hspace{0.5cm}\times \prod_{\ell = 2}^d\frac{\Gamma\left(n_j\alpha_{j\ell} + \zeta_{\ell}\right)}{\left(\sum_{i:\gamma_i = j}r_iy_{i\ell} + \sigma_{\ell}\right)^{n_j\alpha_{j\ell} + \zeta_{\ell}}}
    \end{aligned}
\end{equation*}
The inferential forms for $\beta_{j\ell}$ and its priors are the same as for PG--G.  The 
  normal prior for $\mu$ is conjugate for the log-normal $\bm{\alpha}_j$, and can be sampled
  via a Gibbs step.  Finally, the inverse Wishart prior for $\Sigma$ is again conjugate to
  the log-normal $\bm{\alpha}_j$, implying that it can also be sampled via a Gibbs step.

% EOF