### 2022-06-18 poster_sub.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}

\title{Bayesian non-parametric inference for multivariate peaks-over-threshold models}
\author{Peter Trubey \& Bruno Sans\'{o}}
\date{University of California, Santa Cruz \\ Santa Cruz, CA 95064; USA}

\begin{document}

\maketitle

\begin{abstract}
    \input{./abstract}
\end{abstract}

\paragraph{Keywords:}  Multivariate extremes, Peak over threshold models, Bayesian non-parametric models, Dirichlet process mixtures

\end{document}

% EOF