### 2023-05-27 notation.tex
updates

\documentclass{article}
\usepackage[margin=2cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}

\begin{document}

\begin{description}
\item[$\bm{W},\bm{w}$] - Random, observed variable following multivariate generalized Pareto distribution.
\item[$\bm{Z},\bm{z}$] - Random, observed variable following multivariate standard Pareto distribution.
\item[$\bm{X},\bm{x}$] - Generic random, observed variable in $\mathcal{R}_+^d$.
\item[$\bm{Y},\bm{y}$] - Random, observed variable on $\mathbb{S}_p^{d-1}$.
\item[$\bm{V},\bm{v}$] - Random, observed variable on $\mathbb{S}_{\infty}^{d-1}$; angular component.
\item[$R,r$] - Random, observed variable on $\mathbb{R}_+$; radial component.
\item[$i$] - Iterator over observations.
\item[$j$] - Iterator over mixture components.
\item[$\ell$] - Iterator over dimensions.
\item[$d$] - Number of dimensions.

\end{description}

\end{document}
% EOF