### 2023-05-27 results.tex

\begin{figure*}[htb]
    \centering
    % \caption{Energy score (top row) and Posterior Predictive Loss (bottom row) for various models on simulated data with ascending count of mixture components (indicated by plot heading) and number of dimensions (indicated by horizontal axis).  Gen indicates baseline.\label{fig:simpples}}
    \caption{Average energy score rise over baseline (on $\mathbb{S}_{\infty}^{d-1}$) for various models on simulated data, with ascending count of mixture components (indicated by plot heading) and number of dimensions (indicated by horizontal axis).  Note that pairwise betas is a moment-restricted model.\label{fig:simpples}}
    \includegraphics[height=3in, width = \textwidth]{./images/sim_es_rise}
\end{figure*}

\section{Data illustrations\label{sec:results}} 
\makenote{this title seems weird?}

\subsection{Simulation Study\label{subsec:simulated}}
To evaluate our proposed approach for angular measure estimation 
    we consider simulated datasets on $\mathbb{S}_{\infty}^{d-1}$,
    for values of $d$ ranging between 2 and 32.  We generated each
    dataset as a mixture of multivariate log-normal distributions,
    projected onto $\mathbb{S}_{\infty}^{d-1}$. The generation 
    procedure is detailed in Algorithm~\ref{algo:simulated}.  For each
    model, the data is then projected onto the appropriate $\mathbb{S}_p^{d-1}$
    for model fitting--dirichlet and pairwise betas being fit on 
    $\mathbb{S}_1^{d-1}$, and the projected gamma models being fit on 
    $\mathbb{S}_{10}^{d-1}$.

\begin{algorithm}[h]  % algorithm using algorithm package
        \caption{Simulated Angular Dataset Generation Routine\label{algo:simulated}.
        $\mu_j$, $\Sigma_j$ are the parameters of the mixture component distribution; 
        $\pi$ is the probability vector assigning weight mixture components; $\delta_i$ 
        is the mixture component identifier for each simulated observation.}
        \begin{algorithmic}
        \For{$n_{\text{iter}}$ in $[1,\ldots,10]$}
        \For{$n_{\text{mix}}$ in $[1, 2, 4, 8]$}
            \For{$j$ in $1,\ldots,n_{\text{mix}}$}
                \State Generate $\mu_{j} \sim \mathcal{N}_{32}\left(\bm{0},\bm{I}\right)$
                \State Generate $\Sigma_{j}\sim\mathcal{IW}_{32}\left(70,70 \bm{I}\right)$
            \EndFor
            \State Generate $\pi\sim\text{Dirichlet}(\bm{10})$
            \For{$i$ in $1,\ldots,1000$}
                \State Generate $\delta_i \sim \text{Categorical}(\pi)$
                \State Generate $\bm{X}_i \sim \mathcal{LN}\left(\mu_{[\delta_i]},\Sigma_{[\delta_i]}\right)$
            \EndFor
            \For{$n_{\text{col}}$ in $[2,4,8,16,24,32]$}
                \State Project columns 1 to $n_{\text{col}}$ of ${\bf X}$ onto $\mathcal{S}_{\infty}^{n_{\text{col}} - 1}$ and save.
            \EndFor
        \EndFor
        \EndFor
        \end{algorithmic}
    \end{algorithm}

Figure~\ref{fig:simpples} shows the average rise over baseline in energy score as
    calculated on $\mathbb{S}_{\infty}^{d-1}$ using the kernel metric described
    in Proposition~\ref{prop:rot}, for models trained on simulated data.  After training
    a model, a posterior predictive dataset is generated, and the energy score is
    calculated as a Monte Carlo approximation of Equation~\ref{eq:es}.
    The \emph{baseline} value is the energy score of a new dataset from the same
    generating distribution as the training dataset, evaluated against the training 
    dataset.  We considered a total of four variations of the Dirichlet-process mixture
    of projected gammas model: whether to restrict $\beta_{j\ell} := 1$ as the projected
    \emph{restricted} gamma (DP-PRG) akin to a projected Dirichlet, or allow them to be 
    learned with a gamma prior (DP-PG); and for the shape parameters whether to use a 
    gamma centering distribution (-G) or a log-normal centering distribution (-LN).
    For comparison we also included two other distributions on $\mathbb{S}_{1}^{d-1}$:
    The Dirichlet distribution (with gamma priors), and the pairwise betas distribution
    developed in~\cite{COOLEY2010}.  Note that the pairwise betas model respects
    the moment restriction generally assumed.\makenote{This phrasing needs to be changed.
    The point I am trying to make is that because the moment-restriction assumption is 
    discarded, a moment-restricted model is inadequate, as is borne out in the
    simulation study.}  We observe that the projected gamma models dominate, and within
    that class of model, the choice of centering distribution is irrelevant.  For the
    projected \emph{restricted} gamma models, the choice of a multivariate log-normal 
    centering distribution negatively affects model fidelity.

\subsection{Integrated Vapor Transport\label{subsec:ivt}}
The \emph{integrated vapor transport} (IVT) is a two component vector 
    that tracks the flow of the total water volume in a column of air 
    over a given area \citep{ralph2017}.  IVT is increasingly used in 
    the study of atmospheric rivers because of its direct relationship 
    with orographically induced precipitation \citep{neiman2009water}. 
    Atmospheric rivers (AR) are elongated areas of high 
    local concentration of water vapor in the atmosphere that transport water
    from the tropics around the world. AR can cause extreme 
    precipitation,  something that is usually associated with very large values
    of the IVT magnitude over a whole geographical area. In spite of this, AR
    are fundamental for the water supply of areas like California. Thus the
    importance of understanding the extreme behavior of IVT, including 
    extreme tail dependence. We consider datasets that correspond to IVT
    estimated at two different spatial resolutions. The coarse resolution dataset
    is obtained from the European Centre for Medium-Range Weather Forecasts
    (ECMWF) Interim reanalysis (ERA-Interim) \citep{berrisford2011atmospheric,dee2011era}. 
    The high resolution dataset corresponds to the latest ECMWF 
    observational product, ERA5  \citep{hersbach2020era5}.

\begin{figure}[tb]
    \begin{center}
    \caption{Grid cell locations for ERA-Interim (left) and ERA5 (right).\label{fig:gridlocs}}
    \begin{subfigure}[b]{0.48\textwidth}
        \centering 
        \includegraphics[height=2.5in]{./images/erai_grid}
    \end{subfigure}
    %
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[height=2.5in]{./images/era5_grid}
    \end{subfigure}
    \end{center}
\end{figure}

Our data corresponds to daily average values for the IVT magnitude
    along the coast of California.  The ERA-Interim data used covers the time period 
    1979 through 2014 (37 years) omitting leap days, and eight grid cells that 
    correspond to the coast of California.  The ERA5 data covers the time period 
    1979 through 2019 (42 years) with the same restriction, and  47 grid cells for 
    the coast of California.  This gives us the opportunity to illustrate the 
    performance of our method in multivariate settings of very different dimensions. 
    Figure~\ref{fig:gridlocs} provides a visual representation of the area these grid 
    cells represent.

\begin{algorithm}[htb]
    \caption{Data preprocessing to isolate and transform data exhibiting extreme behavior.  $r_i$
    represents the radial component, and $\bm{v}_i$ the angular component.  The declustering 
    portion is relevant for data correlated in time.\label{algo:processing}}
    \begin{algorithmic}
        \For{$\ell = 1,\ldots,d$}
            \State Set $b_{t,\ell} = \hat{F}_{\ell}^{-1}\left(1 - \frac{1}{t}\right)$.
            \State With $\bm{ x}_{\ell} > b_{t,\ell}$, fit $a_{\ell}$, $\xi_{\ell}$ via MLE according to generalized Pareto likelihood.
        \EndFor
        \For{$i = 1,\ldots,n$}
            \State Define $z_{i,\ell} = \left(1 + \xi_{\ell}\frac{x_{i,\ell} - b_{t,\ell}}{a_{\ell}}\right)_{+}^{1/\xi_{\ell}}$; $\;\;\;$ then $r_i = \pnorm{\bm{ z}_i}{\infty}$, $\;\;\bm{ v}_i = \frac{\bm{ z}_i}{\pnorm{\bm{ z}_i}{\infty}}$
        \EndFor
        \State Subset $\bm{ r},\bm{ v}$ such that $r_i \geq 1$
        \If{declustering}
            \For{$i = 1,\ldots,n$}
                \State If $r_i \geq 1$ and $r_{i-1} \geq 1$, drop the lesser (and associated $v_i$) from data set.
            \EndFor
        \EndIf
    \end{algorithmic}
\end{algorithm}

\begin{figure*}[ht]%{l}{0.45\textwidth}
    \centering
    \caption{Pairwise plots from ERA-Interim data after transformation and projection to ${\mathbb S}_{\infty}^{7}$.  Down the diagonal are marginal kernel densities, with two-dimensional histograms on the off-diagonal.  In those plots, red indicates a higher density.  All data is between 0 and 1.\label{fig:erai_data}}
    \includegraphics[width=.7\linewidth]{./images/data_transformed}
\end{figure*}

Fitting our models to the IVT data requires some pre-processing. First, we subset the data to the rainy
  season, which in California runs roughly from November to March.  Following the approach described 
  in Section \ref{sec:methodology} we estimate the shape and scale parameters of a univariate GP, in each
  dimension, using maximum likelihood. We set the threshold in each dimensions $\ell$ as  
  $b_{t,\ell} = \hat{F}_{\ell}^{-1}(1 - t^{-1})$, where $\hat{F}$ is the empirical CDF and $t=20$, 
  that corresponds to the $95$ percentile. We then use the transformation in Equation
  \eqref{eqn:standardization} to standardize the observations.  Dividing each standardized observation
  by its $\mathcal{L}_{\infty}$ norm, we obtain a projection onto $\mathcal{S}_{\infty}^{d-1}$. As 
  the data correspond to a daily time series, the observations are temporally correlated.  For each
  group of consecutive standardized vectors $z_i$ such  that  $\inorm{z_i} > 1$, we retain only the
  vector with the largest $\mathcal{L}_{\infty}$ norm.  The complete procedure is outlined in
  Algorithm~\ref{algo:processing}.

After subsetting the ERA-Interim data to the rainy season we 
  have $5587$ observations. After the processing and declustering described in
  Algorithm~\ref{algo:processing}, this number reduces to $511$ observations. A pairwise plot 
  of the transformed data after processing and declustering is presented in 
  Figure~\ref{fig:erai_data}.  From this, we note that the marginal densities display strong
  similarities, with a large spike near 0 and a small spike near 1. A value of 1 in a particular 
  axis indicates that the standardized threshold exceedance was largest in that dimension.  
  The off-diagonal plots correspond to pairwise density plots.  We observe that some site pairs, such 
  as $(1,2)$, $(7,8)$, and especially $(4,5)$ have the bulk of their data concentrated in a 
  small arc along the $45^{\circ}$, while other site combinations such as $(3,6)$, $(2,7)$, or
  $(1,8)$ the data is split, favoring one side or the other of the $45^{\circ}$ line. For the 
  ERA5 data, after subsetting we have $6342$ observations, which reduces to $532$ observations 
  after processing and declustering. We fit the PG--G, PRG--G, PG--LN, and PRG--LN models to 
  both datasets.

\begin{table}[htb]
  \centering
  \caption{Energy score criterion from fitted models against the IVT data. 
    Lower is better.\label{tab:dev}}
  \include{./tables/table_dev_updated}
\end{table}

Table~\ref{tab:dev} shows the values of the estimated posterior predictive loss and energy 
    scores for the different models considered. We observe that, as in the case of the simulation 
    study shown in Figures~\ref{fig:simpples}, the preferred model is the projected restricted 
    gamma models, though for the lower-dimensional ERA-Interim data, all models perform comparably.
    Interestingly, this preference for the restricted gamma form is the opposite of the result 
    obtained in the simulation example.

\begin{figure*}[htb]
    \centering
    \caption{Pairwise extremal dependence coefficients for IVT data using the PRG--G model.\label{fig:chi_ij}}
    \includegraphics[width=0.9\textwidth]{./images/chi_ij_c}
\end{figure*}

We consider an exploration of the pairwise extremal dependence using Monte Carlo estimates of the 
  coefficients in  Equation~\ref{eqn:chi_ij}. For this we use samples obtained from the PRG--G model.
  Figure~\ref{fig:chi_ij} provides a graphical analysis of the results. 
  The coefficients achieve values between $0.286$ and $0.759$ for the ERA-Interim data and 
  between $0.181$ and $0.840$ for the ERA5 data.  The greater range in dependence scores observed 
  with the ERA5 data versus ERA-Interim speaks to the greater granularity of the ERA5 data--distance between locations is a strong contributor to the strength of the 
  pairwise asymptotic dependence. The highest coefficients are $0.759$ for 
  locations 4 and 5 in the ERA-Interim data and
  $0.840$ for locations 1 and 2 in the ERA5 data.  Clearly, pairwise asymptotic
  dependence coefficients tell a limited story, as a particular dependence may include
  more than two locations.   We can, however, glean some information from the patterns that
  emerge in two dimensions.  For the ERA-Interim data, we observe a possible cluster 
  between cells 5-8, indicating a strong dependence among these cells.  Analogously, for
  the ERA5 data, we observe three possible groups of locations.

\begin{figure*}[htb]
    \centering
    \caption{Conditional survival curves for selected locations, using ERA-Interim, and PRG-G model,  conditioning on all other dimensions at greater than 90th percentile (fitted)\label{fig:condsurv1d}. The left panel uses original units. Right panel uses standardized units.}
    \includegraphics[width=\linewidth]{./images/condsurv_1d}
\end{figure*}

\begin{figure*}[htb]
    \centering
    \caption{Pairwise conditional survival curves for selected locations, using 
        ERA-Interim, and PRG-G model, conditioning on all other dimensions at greater 
        than 90th percentile (fitted).\label{fig:condsurv2d}}
    \includegraphics[height=4in, width=4in]{./images/condsurv_2d}
\end{figure*}

Figure~\ref{fig:condsurv1d} shows, for the ERA-Interim data under the PRG--G model,  
    the conditional survival curve defined in Equation~\ref{eqn:condsurv}, for one dimension, 
    conditioned on all other dimensions being greater than their (fitted) $90$th percentile. 
    Figure~\ref{fig:condsurv2d} presents the bi-variate conditional survival function,
    conditioning on all other dimensions.  These results illustrate quantitatively how extremal dependence
    affects the shape of the conditional survival curves.  The two top panels represent the joint survival function 
    between grid locations 4 and 5, which are shown in Figure~\ref{fig:chi_ij} to exhibit strong
    extremal dependence.  We observe that the joint survival surface 
    is strongly convex.  The bottom panels represent the joint survival surface between grid locations 1 and 5, 
    which exhibited low extremal dependence.  In this case the shape of the contours tend to be concave, quite different from the shapes observed in the top panels.

Using our proposed scoring criteria, we explored the effect of the choice of $p$ on the final results. Using the simulated data, generated
    from a mixture of projected Gammas, we were unable to observe sizeable differences in the scores for $p$ 
    ranging between 1 and 15.   However, 
    for the IVT data, we observed a drop in the energy score associated 
    with higher $p$, with diminishing effect as $p$ increased.  
    We observed no significant differences in the performance
    of the model that uses $p=10$, which corresponds to the analysis 
    presented, relative to the one that uses $p=15$.
% EOF