### 2023-06-20 conclusion.tex

\section{Conclusion\label{sec:conclusion}}
In this paper, we have built upon the definition of the multivariate Pareto 
    described in \cite{ferreira2014} to establish a useful representation of 
    its dependence structure, through the distribution of its angular component,
    which is supported on the positive orthant of the unit hypersphere under 
    the $\mathcal{L}_{\infty}$ norm.  Due to the inherent difficulty of 
    obtaining the likelihood of distributions with support on ${\mathbb S}^{d-1}_\infty$ 
    our method transforms the data to ${\mathbb S}_{p}^{d-1}$, fits then using mixtures of 
    products of independent gammas, then transforms the predictions back to 
    ${\mathbb S}^{d-1}_\infty$. As ${\mathbb S}_{p}^{d-1}$ converges to 
    ${\mathbb S}_{\infty}^{d-1}$ as  $p\to\infty$, we expect the proposed 
    resampling to be efficient for large enough $p$. In fact, our exploration 
    of the simulated and real data indicates that the procedure is robust to 
    the choice of moderately large values of $p$. 
    Our method includes two inferential steps. The first consists of the
    estimation of the marginal Pareto distributions. The second consists 
    of the estimation of the angular density. Parameter uncertainty incurred in
    the former is not propagated to the latter. Conceptually, and integrated approach 
    that accounts for all the estimation uncertainty is conceivable. Unfortunately, 
    this leads to posterior distributions with complex data dependent restrictions 
    that  are very difficult to explore, especially in large dimensional setting. In 
    fact, our attempts to fit a simple parametric model for the marginals and the
    angular measure jointly in several dimensions were not successful.

In this paper we have focused on a particular representation of the multivariate
    Pareto distribution for PoT inference on extreme values. To this end, our model 
    provides a computationally efficient and flexible approach. An interesting extension 
    of the proposed model is to consider regressions of extreme value responses, due
    to extreme value inputs following the ideas in  \cite{carvalho2022}. This will 
    produce PoT based Bayesian non-parametric extreme value regression models. More 
    generally,  models that allow for covariate-dependent extremal dependence 
    \citep{mhalla2019} could be considered. In addition, we notice that our approach 
    is based on flexibly  modeling angular distributions for any $p$-norm. As such, 
    it can be applied to other problems focused on high dimensional directional 
    statistics  constrained  to a cone of directions.

We explored the behavior of the proposed model using simulated data from
    mixtures of projected Gammas with varying degrees of complexity---both 
    in terms of dimensionality and in number of mixture components.  From this,
    we observed that the additional flexibility offered by varying the $\beta$
    term---the rate parameter of the gamma distribution---can offer additional
    model fidelity.  \makenote{However,} this result is not borne out in our 
    real data example, for which in higher dimensions the simpler model is 
    strongly preferred.  Additionally, we observe
    that the choice of the log-normal centering distribution and additional
    flexibility it offers did not translate to additional model fidelity.  
    \makenote{This result is borne out in the real data, where it performed at best 
    comparably to the gamma model.}  
    % This
    % is borne out in the real data, where on the ERA--Interim dataset it performed
    % only comparably, but on ERA--5 it performed significantly worse.
    % {\bf I am not sure that this paragraph is appropriate anymore. I think that
    % the different mixture models considered are pretty much equivalent, so we should
    % favor the one that is easiest and most robust to fit.}

The computations in this paper were performed on a desktop computer with an 
    AMD Ryzen 5950X processor. The program is largely single-threaded, so 
    computation time is not dependent on available core count.  In each case, 
    we run the MCMC chain for 50\,000 iterations, with a burn-in of 40\,000 
    samples.  Fitting the PG--G model on the ERA5 dataset took approximately 
    15 minutes.  Work is in progress to optimize the code, and explore 
    parallelization where possible.  We are also exploring alternative 
    computational approaches that will make it feasible to tackle very high 
    dimensional problems, such as variational Bayes. In fact, to elaborate on 
    the study of IVT, there is a need to consider several hundreds, if not thousands, 
    of grid cells over the Pacific Ocean in order to obtain a good description of 
    atmospheric  events responsible for large storm activity over California.

% §6 calls for a view of the authors on what 
%     could be related exercises—or even extensions of this research and future 
%     work. Using the proposed Bayesian multivariate PoT construction for 
%     devising a regression model for an an extreme value response on an extreme 
%     value covariate is one example of a natural follow-up; see 
%     \cite{carvalho2022} for related ideas. Conditional versions of the angular 
%     measure (e.g. \cite{carvalho2016}, \cite{castro2018}, \cite{escobar2018}, 
%     \cite{mhalla2019}) would lead to a conditional multivariate 
%     peaks-over-threshold model, where dependence between exceedances could 
%     change along with a covariate. Also, perhaps the here proposed DPM of 
%     projected gammas—as well as predictor-dependent versions of it—could be of 
%     interest in other settings beyond angular measures?

% EOF