### 2024-02-01 abstract_sbss.tex
changes

We consider a constructive definition of the multivariate Pareto that factorizes the 
random vector into a radial component and an independent angular component; The former
following a univariate Pareto distribution, and the latter defined on the surface of the 
positive orthant of the infinity norm unit hypercube.   We propose a method 
for inferring the distribution of the angular component, by identifying its support 
as the limit of the positive orthant of the unit $p$--norm spheres, and introduce a 
projected gamma family of distributions defined through the normalization of a vector of 
independent random gammas to the space.   This serves in constructing a flexible family of
distributions obtained as a Dirichlet process mixture of projected gammas.  For model 
assessment, we discuss scoring methods appropriate to distributions on the unit hypercube.  
In particular, working with the energy score criterion, we develop a kernel metric that 
produces a proper scoring rule, and present 
a simulation study to compare different modeling choices using the proposed metric.    
Using our approach, we describe the dependence structure of extreme values in the
integrated vapor transport (IVT), data describing the flow of atmospheric moisture 
along the coast of California.  We find clear but heterogeneous geographical dependence.