### 2024-02-01 main_sbss.tex

\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{soul}
\usepackage{bm}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{setspace}
\linespread{1.8}

\usepackage{multirow}%
\usepackage{mathrsfs}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{anysize}
\usepackage[margin=1.75cm]{geometry}

\newtheorem{prop}{Proposition}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\bigmid}{\,\middle\vert\,}
\newcommand{\bruno}[1]{\textcolor{red}{#1}} % Bruno edits - red
\newcommand{\makenote}[1]{\textcolor{teal}{#1}} % Peter edits - teal

\newcommand{\findcite}{\textcolor{red}{[Find Citation]}}
\newcommand{\needcite}{\findcite}
\newcommand{\Chi}{\mbox{\Large$\chi$}}

\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\inorm}[1]{\norm{#1}_{\infty}}
\newcommand{\pnorm}[2]{\norm{#1}_{#2}}
\newcommand{\prob}[1]{\text{Pr}\left[#1\right]}
\newcommand{\expect}[1]{\text{E}\left[#1\right]}

\raggedbottom

\title{Bayesian Non-parametric Inference for Multivariate\\ Peaks-over-threshold Models}

% \author{Peter Trubey \& Bruno Sans\'{o}\\ Department of Statistics, UC Santa Cruz}
\author{}
\date{}

\begin{document}

\maketitle

\abstract{\input{./abstract_sbss}}
% \keywords{Multivariate extremes, Peak over threshold models,
%                 Bayesian non-parametric models, Dirichlet process mixtures}

\maketitle

%--------content
\input{./introduction}

\input{./methodology}

\input{./evaluation}

\input{./results}

\input{./conclusion}

\section*{Acknowledgements}
This material is based upon work supported by the National Science Foundation
    under Grant No. SES-2050012 and DMS-2153277.

\bibliographystyle{plainnat}
\bibliography{./refs}

\end{document}

% EOF