### 2024-02-20 cover.tex
entropy submission RC1

\documentclass[letterpaper,11pt]{letter}

\name{Peter Trubey}
\signature{Peter Trubey}
\address{Peter Trubey \\ Department of Statistics \\ University of California, Santa Cruz \\ USA}

\begin{document}%

\begin{letter}{Editorial Board \\ \emph{Entropy}}

\opening{Dear Editorial Board,}

We submit to you our manuscript for inclusion in the special issue of \emph{Entropy}.  It details a method of 
inference for the dependence structure of extremes in a multivariate peaks over thresholds model, based on a 
mixture of projections of vectors of gamma variables.  Included are an appropriate model selection 
criterion, and applications of the fitted model.  A simulation study is performed to explore properties of the 
distribution, and the model then applied to atmospheric data.

We confirm that neither the manuscript nor any parts of its content are currently under consideration or published in another journal.

All authors have approved the manuscript and agree with its submission to \emph{Entropy}.

Sincerely,

\closing{~}

\end{letter}
\end{document}

% EOF