### 2024-04-10 abstract.tex
push for mdpi resubmission

We consider a constructive definition of the multivariate Pareto that factorizes the 
random vector into a radial component and an independent angular component; The former
following a univariate Pareto distribution, and the latter defined on the surface of the 
positive orthant of the infinity norm unit hypercube.  In this paper, we  propose a method 
for inferring the distribution of the angular component.  We identify its support 
as the limit of the positive orthant of the unit $p$-norm spheres, and introduce a 
projected gamma family of distributions defined through the normalization a vector of 
independent gamma random variables to the $p$-norm sphere.  This family serves as a 
building block for a flexible family of distributions obtained as a Dirichlet process 
mixture of projected gammas.  For model assessment, we discuss scoring methods appropriate 
to distributions on the unit hypercube.  In particular, working with the energy score
criterion, we develop a kernel metric that produces a  proper scoring rule, and present 
a simulation study to compare  different modeling choices using the proposed scoring rules.  
Finally we apply our  approach to describe the dependence structure of the extreme values of 
the magnitude  of the integrated vapor transport (IVT), data describing the flow 
of atmospheric moisture along the coast of California for the years of 1979 
through 2020.  We find clear but heterogeneous geographical dependence.